---
title: "spotify genre predictions"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r, r-setup, warning=F, message=F, echo=T, results="markup"}
library(MASS)
library(httr)
library(jsonlite)
library(knitr)
library(car)
library(tidyverse)
```

## Initialization

Setup the API credentials to scrape our data from Spotify.

``` {r, api-setup} 
myId <- "320f39b48a7946319c7040d2f877e257" 
mySecret <- "5487fde2cc5349ad902d58a908752995"

response <- POST(
  'https://accounts.spotify.com/api/token',
  accept_json(),
  authenticate(myId,mySecret),
  body = list(grant_type = "client_credentials"),
  encode = "form"
)

myToken <- content(response)$access_token

myHeader <- paste0('Bearer ', myToken)
```

Write a function to scrape song ID's from playlists.

```{r, trackScrape function}
# for getting the tracks from playlist
trackScrape <- function(playlist, genre="NA") {
  output <- NULL
  URI <- paste0("https://api.spotify.com/v1/playlists/", playlist, "/tracks")
  dat <- content(GET(URI, add_headers(Authorization = myHeader)))
  for (i in 1:dat$total){
    id <- dat$items[[i]]$track$id
    title <- dat$items[[i]]$track$name
    artist <- dat$items[[i]]$track$artists[[1]]$name
    output <- rbind(output,c(id,title,artist,genre))
  }
  colnames(output)=c("ID","Title","Artist","Genre")
  output
} 
```

Write a function to scrape song features from ID's.

```{r, featureScrape function}
# for getting the features of tracks
featureScrape <- function(tracks) {
  output <- NULL
  URI <- paste0("https://api.spotify.com/v1/audio-features/?ids=", paste(tracks[,1],collapse=","))
  dat <- content(GET(URI, add_headers(Authorization = myHeader)))$audio_features
  for (i in 1:nrow(tracks)){
    output <- rbind(output,as.data.frame(dat[[i]]))
  }
# only want columns 1 2 4 6:8 10 11 17  
  output <- cbind(tracks[,2:3],output[,c(1,2,4,6:8,10,11,17)],tracks[,4])
  colnames(output)[12] <- "Genre"
  output
}
```

Combine the two functions so we just feed it playlists to get the features of each song.

```{r, myScrape function}
# combine the two for fun times
myScrape <- function(playlist, genre="NA") {
  output <- NULL
  for (i in 1:length(playlist)){
    Tracks <- trackScrape(playlist[i], genre[i])
    Features <- featureScrape(Tracks)
    output <- rbind(output, Features)
  }
  output
}
```

Scrape the data using our functions and write them to files for ease of use in the future.

```{r, dataset creation, echo = F, eval = F}
pop.list <- "6lXAHzJaMNTmEhWm3Wk5Qr"
kpop.list <- "6K5LCPJtDmddD5pTNp5KRD"
old.list <- "7I6AKVkiaZS8Ih8KE50dvB"
rock.list <- "1LNujkCkKz4VSN9wc7h6Au"
country.list <- "7mpQKYCqlDc0W1PFPqfXdV"
class.list <- "6aZe42NMzR0PEwX82X0DUa"
rap.list <- "2bKotyC4JNjEnhIY6nKvuw"
edm.list <- "6XeEpxLG2c69yi1R9moDKv"

# setup for the function
full.list <- c(pop.list, kpop.list, old.list, rock.list, country.list, class.list, rap.list, edm.list)
full.genre <- c("Pop","K-Pop","Oldies","Rock","Country","Classical","Rap","EDM")

# scrape the data
dat <- myScrape(full.list, full.genre)
dat$duration_ms <- dat$duration_ms/1000 # turn from ms to seconds. also forces class to dbl for QOL
names(dat)[names(dat)=="duration_ms"] <- "duration" # no longer in ms
dat$Title <- str_trunc(as.character(dat$Title), 30) # truncate song names...
dat$Artist <- str_trunc(as.character(dat$Artist), 20) # ...and artists

# write the data so we don't have to scrape in the future.
write.table(dat, file = "project.dat", row.names = F)
```

Read and subset the data into 3 groups:

1. Partial data which holds all the songs but does not specify the subgroups for Pop
2. Pop data which holds only songs from the Pop subgroups
3. Full data which holds all the songs and does specify the subgroups for Pop

```{r, read and subset our data}
dat0 <- as.tibble(read.table("project.dat",h=T)) # read as tibble for use of dplyr library

# subset our data
artist.dat <- dat0[,1:2] # put songs and artists in separate variable for QOL
full.dat <- dat0[,-c(1:2)] # remove songs and artists from data for QOL
partial.dat <- full.dat # partial.dat will be with all Pop subgenres combined
partial.dat$Genre[partial.dat$Genre %in% c("Pop","K-Pop","Oldies")] <- "Pop" # combine the Pop subgroups
pop.dat <- full.dat[full.dat$Genre %in% c("Pop","K-Pop","Oldies"),] # data using only the Pop subgenres
```
## EDA

In order to get a correlation matrix that sorts rows by distance it takes some work:

```{r, melter function and outputs}
melter <- function(dat) {
  # reordering equation gathered online from here...
  dd <- as.dist((1-dat)/2)
  hc <- hclust(dd)
  dat <- dat[hc$order, hc$order]
  # ... until here
  dat[lower.tri(dat)] <- NA
  dimension <- nrow(dat)
  datnames <- rownames(dat)
  output <- data.frame(v1 = factor(datnames, ordered = T, levels = datnames), v2 = factor(datnames, ordered = T, levels = datnames), cor = 1:dimension)[0,] # ordered bc of ggplot's geom_tile function
  for (i in 1:nrow(dat)){
    for (j in i:nrow(dat)){
      output[nrow(output)+1,] <- c(datnames[i], datnames[j], as.double(dat[i,j]))
    }
  }
  output[,3] <- as.double(output[,3])
  output
}

cor.dat <- melter(cor(partial.dat[,-10]))

ggplot(cor.dat , aes(x = v1, y = v2, fill = cor, label = round(cor,3))) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1,1), name = "Correlation") +
  geom_text(aes(size = abs(cor)), show.legend = F) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.direction = "horizontal",
    legend.position = c(.8,.1))
```

```{r, comparisons amongst genres}
aggregate(full.dat[,-10], by=full.dat['Genre'], FUN=mean)
aggregate(partial.dat[,-10], by=partial.dat['Genre'], FUN=mean)
aggregate(pop.dat[,-10], by=pop.dat['Genre'], FUN=mean)
```

## LDA

basic lda work

```{r, lda stats}
full.lda <- lda(Genre~.,dat=full.dat)
partial.lda <- lda(Genre~.,dat=partial.dat)
kable(round(full.lda$means,4))

ggplot(data = as.data.frame(predict(full.lda)$x),mapping=aes(x=LD1, y=LD2)) +
  geom_point(mapping=aes(color=full.dat$Genre)) +
  scale_color_discrete(name="Genre") +
  ggtitle("Full Model LDA Predictions")

ggplot(data = as.data.frame(predict(partial.lda)$x),mapping=aes(x=LD1, y=LD2)) +
  geom_point(mapping=aes(color=partial.dat$Genre)) +
  scale_color_discrete(name="Genre") +
  ggtitle("Partial Model LDA Predictions")
```

check the accuracy of our models:

```{r, checking accuracies}
confusion.full <- table(Actual=full.dat$Genre,Predicted=predict(full.lda)$class)
confusion.partial <- table(Actual=partial.dat$Genre,Predicted=predict(partial.lda)$class)[-c(4:5),-c(4:5)]
cbind(confusion.full,Acc=round(diag(confusion.full)/rowSums(confusion.full),3))
cbind(confusion.partial,Acc=round(diag(confusion.partial)/rowSums(confusion.partial),3))
round(c(sum(diag(confusion.full))/nrow(full.dat),sum(diag(confusion.partial))/nrow(partial.dat)),3) # accuracy rate, full vs partial

```

i want to try something similar to a discrepancy/distance chart for the full model to create a distance map thing

this probably doesn't really make sense but do it for fun

```{r, try making a distance using confusion + t(confusion) for full}
par(mfrow=c(1,2))
distance.full <- sqrt(confusion.full + t(confusion.full))
distance.full <- (distance.full-diag(distance.full))/diag(distance.full)
fit.full <- cmdscale(distance.full,eig=T,k=2)
plot(fit.full$points,pch=0, type="n", xlim=c(-1,1),ylim=c(-1,1),main="Full Model")
text(fit.full$points,row.names(fit.full$points),offset=4)

distance.partial <- sqrt(confusion.partial + t(confusion.partial))
distance.partial <- (distance.partial-diag(distance.partial))/diag(distance.partial)
fit.partial <- cmdscale(distance.partial,eig=T,k=2)
plot(fit.partial$points,pch=0, type="n", xlim=c(-1,1),ylim=c(-1,1),main="Partial Model")
text(fit.partial$points,row.names(fit.partial$points),offset=4)
```

want to use just subset our data into training and testing data for predictions. decide not to use priors: the distribution i created is not indicative at all of the actual distribution (should this model ever be used with new songs)

```{r, LDA function}
myLDA <- function(data, n = 30, seed=8675309) {
  set.seed(seed)
  
  testindex <- sample(nrow(data),n)
  testdat <- data[testindex,]
  dat <- data[-testindex,]
  
  output <- lda(Genre~.,dat)
  
  output <- cbind(artist.dat[testindex,],predict(output, newdata=testdat)$class,data$Genre[testindex])
  output <- cbind(output,output[,3] == output[,4])
  colnames(output)[3:5] <- c("Predicted", "Actual", "Accuracy")
  output
}
```

using our function, want to compare if full better than partial. first make a function for easy comparing

```{r, LDA comparing function}
LDAcompare <- function(lda1,lda2){ # assumes same seed and underlying observations for lda1 and lda2
  output <- cbind(lda1[,1:4],lda2[,3:4],Accuracy=NA)
  val1 <- as.numeric(lda1[,5])
  val2 <- as.numeric(lda2[,5])*2
  output$Accuracy[val1+val2==0] <- "None"
  output$Accuracy[val1+val2==1] <- "LDA1"
  output$Accuracy[val1+val2==2] <- "LDA2"
  output$Accuracy[val1+val2==3] <- "Both"
  colnames(output)[c(3,5)] <- c("LDA1","LDA2")
  output
}
```

```{r, LDA random testing}
myseed=sample(1:10000,1) # so we can have the same test data for both partial and full models

partial.lda <- myLDA(partial.dat, seed=myseed)
full.lda <- myLDA(full.dat, seed=myseed)
pop.lda <- myLDA(pop.dat, n=10, seed=myseed)

mysums <- c(sum(partial.lda$Accuracy),sum(full.lda$Accuracy))
mypcts <- round(mysums/c(nrow(partial.lda),nrow(full.lda)),4)
mytable <- rbind(no.Correct=mysums,pct.Correct=mypcts)

kable(LDAcompare(partial.lda,full.lda))
mytable
```

write a similar testing function for just the partial and full model but for an even distribution of genres in the sample

```{r, LDA2 function}
myLDA2 <- function(data, triplet=1, seed=8675309) { #take out multiples of 3 due to 3 subsets of Pop
  set.seed(seed)
  
  genres <- levels(full.dat$Genre)
  testindex <- NULL
  
  for (i in 1:length(genres)){
    if (genres[i] %in% c("Pop","K-Pop","Oldies")){
      testindex <- c(testindex,sample(which(full.dat$Genre==genres[i]),triplet))
    } else {
      testindex <- c(testindex,sample(which(full.dat$Genre==genres[i]),triplet*3))
    }
  }

  testdat <- data[testindex,]
  dat <- data[-testindex,]
  
  output <- lda(Genre~.,dat)
  
  output <- cbind(artist.dat[testindex,],predict(output, newdata=testdat)$class,data$Genre[testindex])
  output <- cbind(output,output[,3] == output[,4])
  colnames(output)[3:5] <- c("Predicted", "Actual", "Accuracy")
  output
}
```

```{r, LDA2 random testing}
myseed=sample(1:10000,1) # so we can have the same test data for both partial and full models

partial.lda <- myLDA2(partial.dat, triplet=2, seed=myseed) # make sure 'triplet' param same for both
full.lda <- myLDA2(full.dat, triplet=2, seed=myseed)

mysums <- c(sum(partial.lda$Accuracy),sum(full.lda$Accuracy))
mypcts <- round(mysums/c(nrow(partial.lda),nrow(full.lda)),4)
mytable <- rbind(no.Correct=mysums,pct.Correct=mypcts)

kable(LDAcompare(partial.lda,full.lda))
mytable
```

some interesting observations:
 1. ironically the partial model is more accurate in predicting Pop. this is because there are 1/3 as many options for the partial model. usually when the partial model were to be correct, the full model just happens to guess the wrong subgenre of Pop; in a few cases the full model will guess Rock or EDM
    + eg "P.Y.T. (Pretty Young Thing)" by Michael Jackson (Pop/Oldies) is classified correctly as Pop by the partial model, but incorrectly as K-Pop by the full model.
    + eg "Rock Your Body" by Justin Timberlake (Pop/Contemporary) is correctly classified as Pop by the partial model, but incorrectly as Rock by the full model.
 2a. however, the full model is much more likely to disagree with the partial model when it incorrectly predicts a song to be Pop
    + eg "Hot In Herre" by Nelly (Rap) is incorrectly classified as Pop by the partial model, but correctly classified by the full model.
    + eg "Harder Better Faster Stronger" by Daft Punk (EDM) is incorrectly classified as Pop by the partial model, but incorrectly classified as Rap by the full model.
 2b. on the rare chance both models incorrectly predict Pop, the expanded model seems 'less' wrong because its more specific
    + eg "I was Made for Lovin' You" by KISS (Rock) is classified as Pop by the partial model. seems like a bit of a stretch. However, the full model classifies as Oldies and suddenly it seems less wrong.
 3. funnily enough the full model will also disagree when the partial model if it determines a song to not be Pop.
    + eg "Bartender" by T-Pain (Rap) is correctly classified by the partial model, but incorrectly classified as Pop by the full model.
 4. the full model also seems to be more likely to predict country music is EDM
    + eg "There Goes My Life" by Kenny Chesney (Country) classified Rock by partial, EDM by full
    + eg "Before He Chears" by Carrie Underwood (Country) classified Country by partial, EDM by full
    
conclusions:
by sheer score, the partial model will win because things are more general
if we were to consider "distance" i think the full model might be slightly better, other than the EDM thing
the full model is definitely a snob for pop music

well would we have better pop subgenre selection if we were to ONLY train using pop data?

```{r, LDA3 function}
myLDA3 <- function(data, triplet=1, seed=8675309) { #take out multiples of 3 due to 3 subsets of Pop
  set.seed(seed)
  
  genres <- c("Pop","K-Pop","Oldies")
  testindex <- NULL
  
  for (i in 1:length(genres)){
      testindex <- c(testindex,sample(which(full.dat$Genre==genres[i]),triplet))
  }

  testdat <- data[testindex,]
  dat <- data[-testindex,]
  
  output <- lda(Genre~.,dat)
  
  output <- cbind(artist.dat[testindex,],predict(output, newdata=testdat)$class,data$Genre[testindex])
  output <- cbind(output,output[,3] == output[,4])
  colnames(output)[3:5] <- c("Predicted", "Actual", "Accuracy")
  output
}
```

```{r, LDA3 triple compare function}
LDA3compare <- function(lda1,lda2,lda3){ # assumes same seed and underlying observations
  output <- cbind(lda1[,1:3],lda2[,3],lda3[,3:4],Accuracy=NA)
  val1 <- as.numeric(lda1[,5])+1
  val2 <- as.numeric(lda2[,5])+3
  val3 <- as.numeric(lda3[,5])+5
  output$Accuracy[val1*val2*val3==15] <- "None"
  output$Accuracy[val1*val2*val3==30] <- "1 Only"
  output$Accuracy[val1*val2*val3==20] <- "2 Only"
  output$Accuracy[val1*val2*val3==18] <- "3 Only"
  output$Accuracy[val1*val2*val3==40] <- "1 & 2"
  output$Accuracy[val1*val2*val3==24] <- "2 & 3"
  output$Accuracy[val1*val2*val3==36] <- "1 & 3"
  output$Accuracy[val1*val2*val3==48] <- "All"
  colnames(output)[c(3,4,5)] <- c("Partial1","Full2","Pop3")
  output
}
```

```{r, LDA3 random testing}
myseed=sample(1:10000,1) # so we can have the same test data for both partial and full models

partial.lda <- myLDA3(partial.dat, triplet=2, seed=myseed) # make sure 'triplet' param same for both
full.lda <- myLDA3(full.dat, triplet=2, seed=myseed)
pop.lda <- myLDA3(pop.dat, triplet=2, seed=myseed)

mysums <- c(sum(partial.lda$Accuracy),sum(full.lda$Accuracy),sum(pop.lda$Accuracy))
mypcts <- round(mysums/c(nrow(partial.lda),nrow(full.lda),nrow(pop.lda)),4)
mytable <- rbind(no.Correct=mysums,pct.Correct=mypcts)

kable(LDA3compare(partial.lda,full.lda,pop.lda))
mytable
```

hypothesis: K-Pop is ruining everything. maybe its not even a "real" genre. check lda scalings to verify

```{r, lda scalings}
pop.manova <- manova(as.matrix(pop.dat[,-10])~pop.dat$Genre)
summary(pop.manova) # for all 9 variables, it apparently is
summary(aov(as.matrix(pop.dat[,-10])~pop.dat$Genre))
# no significant differences @ 5%: danceability, acousticness, tempo, duration

# retry this for all 3 combinations of 2 subgenres of pop?
pop.manova1 <- manova(as.matrix(pop.dat[pop.dat$Genre %in% c("Pop","K-Pop"),-10])~as.matrix(pop.dat[pop.dat$Genre %in% c("Pop","K-Pop"),10]))
pop.manova2 <- manova(as.matrix(pop.dat[pop.dat$Genre %in% c("Pop","Oldies"),-10])~as.matrix(pop.dat[pop.dat$Genre %in% c("Pop","Oldies"),10]))
pop.manova3 <- manova(as.matrix(pop.dat[pop.dat$Genre %in% c("Oldies","K-Pop"),-10])~as.matrix(pop.dat[pop.dat$Genre %in% c("Oldies","K-Pop"),10]))

summary(pop.manova1) # Pop vs K-Pop p-value: 0.01387
summary(pop.manova2) # Pop vs Oldies p-value: 1.784e-06
summary(pop.manova3) # Oldies vs K-Pop p-value: 0.002293
```

```{r, LDA scalings for pop only}
pop.lda <- lda(Genre~., dat=pop.dat)

ggplot(data = as.data.frame(predict(pop.lda)$x),mapping=aes(x=LD1, y=LD2)) +
  geom_point(mapping=aes(color=pop.dat$Genre)) +
  scale_color_discrete(name="Genre") +
  ggtitle("Pop Model LDA Predictions")
```

K-Pop definitely seems to be the weak link here

```{r, prelim remove kpop}
nok.full.dat <- full.dat[full.dat$Genre!='K-Pop',]
nok.partial.dat <- nok.full.dat
nok.partial.dat$Genre[nok.full.dat$Genre=='Oldies'] <- 'Pop'

nok.full.lda <- lda(Genre~.,nok.full.dat)
nok.partial.lda <- lda(Genre~.,nok.partial.dat)

confusion.nok.full <- table(Actual=nok.full.dat$Genre,Predicted=predict(nok.full.lda)$class)[-4,-4]
confusion.nok.partial <- table(Actual=nok.partial.dat$Genre,Predicted=predict(nok.partial.lda)$class)[-c(4:5),-c(4:5)]

cbind(confusion.full,Acc=round(diag(confusion.full)/rowSums(confusion.full),3))
cbind(confusion.nok.full,Acc=round(diag(confusion.nok.full)/rowSums(confusion.nok.full),3))
round(c(sum(diag(confusion.full))/nrow(full.dat),sum(diag(confusion.nok.full))/nrow(nok.full.dat)),3)  # accuracy rate, full vs full w/o k-pop

cbind(confusion.partial,Acc=round(diag(confusion.partial)/rowSums(confusion.partial),3))
cbind(confusion.nok.partial,Acc=round(diag(confusion.nok.partial)/rowSums(confusion.nok.partial),3))
round(c(sum(diag(confusion.partial))/nrow(partial.dat),sum(diag(confusion.nok.partial))/nrow(nok.partial.dat)),3)  # accuracy rate, partial vs partial w/o k-pop
```

k-pop isn't a real genre tell your friends

## FA
```{r, fa}
fa.full <- factanal(full.dat[,-c(10)],factors=4,scores="regression",rotation="promax")
```

write a function to tell me what i want so i can scroll less:

```{r, fa cumulative variance and proportional variance calculation}
favar <- function(dat, f=1, rot='promax') {
  x <- factanal(dat,factors=f, scores="regression", rotation=rot)
  chart <- data.frame()
  for (i in 1:f){
    chart[1,i] <- round(sum(x$loadings[,i]^2)/nrow(x$loadings),3)
    if (i > 1){chart[2,i] <- chart[2,i-1] + chart[1,i]}
    else {chart[2,i] <- chart[1,i] }
  }
  rownames(chart) <- c("Proportional Var", "Cumulative Var")
  chart <- list(Variance = chart, PVal = as.numeric(x$PVAL))
  chart
}

favar(full.dat[,-10], f=4)
```

remove 1 var at a time to compare var and pval just to see:

```{r, fa removing 1 var at a time to compare}
faremove <- function(dat, i=10, f=4) {
favar(dat[,-i],f)
}

faselect <- function(dat){
varchart <- matrix(0,nrow=10,ncol=4)
pvalchart <- matrix(0,nrow=10,ncol=4)

for(k in 1:4){
  varchart[10,k] <-faremove(dat,i=10,f=k)$Variance[2,k]
  pvalchart[10,k] <- round(faremove(dat,i=10,f=k)$PVal,3)
}

for (j in 1:9){
  for (k in 1:4){
    varchart[j,k] <- faremove(dat,i=c(j,10),f=k)$Variance[2,k]
    pvalchart[j,k] <- round(faremove(dat,i=c(j,10),f=k)$PVal,3)
  }
}
list("CumVar"=varchart,"PVal"=pvalchart)
}

```
